---
title: "Simple R Example to Import JSON"
date: "2024-10-24"
output: html
execute:
  echo: true
---

# This is a very basic quarto document for sharing your code


This is a simple example of how to import a json.
JavaScript Object Notation, or JSON, is a semi-structured format often used by web applications and REST APIs to exchange data.
To run this, pacman and rmarkdown will need to be installed.

```{r, prep-for-example}

# Install some packages
if (!require("pacman")) install.packages("pacman")
if (!require("readr")) install.packages("readr")


# Load necessary library
pacman::p_load(pacman, tidyverse, jsonlite, rmarkdown)
```

Fetch some data by calling an API

```{r, get-some-data}

# Call an API to get JSON data
data <- "https://ghoapi.azureedge.net/api/Dimension" %>%
  fromJSON() %>%
  print()

```

::: {.callout-note}
This API call only shows available dimensions. 
:::

That is how to fetch and read a JSON file.

```{r, reformat-data-for-vis}
# Re-format the content
data %>% 
  toJSON(pretty = T) %>%
  print()

```

::: {.callout-tip}
You can create callout-tips! Just use the `callout-tip` class.
:::


