---
title: "DHS family planning Analysis"
format: html
---

### Libraries needed for ODBC / Spark connection to Databricks

```{r setup, message=FALSE, warning=FALSE}
# Load required libraries 

# Helper function to install required packages if not already installed
install_if_missing <- function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
    library(pkg, character.only = TRUE)
  }
}

# Install and load necessary packages
packages <- c("DBI", "odbc", "haven", "dplyr", "dbplyr", "janitor", "purrr", "sparklyr", "pysparklyr", "reticulate", "usethis")
invisible(lapply(packages, install_if_missing))

```

### connect to databricks

-   make sure you have downloaded and installed databricks ODBC driver <https://www.databricks.com/spark/odbc-drivers-download>

```{r}
# Make sure driver is installed
print((odbc::odbcListDrivers() %>% filter(grepl("Spark",value )))$name)

# Set these variables or use usethis::edit_r_environ() to store them
# Sys.setenv(DATABRICKS_HOST = "")
# Sys.setenv(DATABRICKS_HTTP_PATH = "")
# Sys.setenv(DATABRICKS_TOKEN = "") 
# Sys.setenv(DATABRICKS_CLUSTER_ID = "")

server_hostname <- Sys.getenv("DATABRICKS_HOST")
http_path       <- Sys.getenv("DATABRICKS_HTTP_PATH")
access_token    <- Sys.getenv("DATABRICKS_TOKEN")

con <- dbConnect(odbc::odbc(),
  Driver       = "Simba Spark ODBC Driver",
  Host         = server_hostname,
  Port         = 443,
  HTTPPath     = http_path,
  AuthMech     = 3,  # Token-based authentication
  UID          = "token",
  PWD          = access_token,
  SSL          = 1,
  ThriftTransport = 2  # Required for HTTPS
)

```

### Query Volumes available (where stata files are located)

```{r}
# Run SQL query to get volumes from information_schema
volume_info <- DBI::dbGetQuery(con, "
  SELECT volume_catalog, volume_schema, volume_name, storage_location
  FROM system.information_schema.volumes
  WHERE volume_catalog = 'dhs_indicator_dev'
  AND volume_schema = 'variables'
")

# View results
# print(volume_info)

```

### Helper function to recursively list file

```{r}
list_dta_files <- function(con, root_path, filename_prefix = NULL) {
  results <- data.frame()
  paths_to_check <- c(root_path)

  while (length(paths_to_check) > 0) {
    current_path <- paths_to_check[1]
    paths_to_check <- paths_to_check[-1]

    # LIST current path
    query <- paste0("LIST '", current_path, "'")
    res <- tryCatch(DBI::dbGetQuery(con, query), error = function(e) NULL)
    if (is.null(res) || nrow(res) == 0) next

    # Heuristic: entries with a dot in name are likely files
    is_file <- grepl("\\.", res$name)

    # Queue subdirectories (no dot in name)
    paths_to_check <- c(paths_to_check, file.path(current_path, res$name[!is_file]))

    # Filter to .dta files only
    dta_hits <- res[is_file & grepl("\\.dta$", res$name, ignore.case = TRUE), ]

    # Optional: filter by filename prefix
    if (!is.null(filename_prefix)) {
      prefix_pattern <- paste0("^(", paste0(filename_prefix, collapse = "|"), ")")
      dta_hits <- dta_hits[grepl(prefix_pattern, basename(dta_hits$name), ignore.case = TRUE), ]
    }

    results <- bind_rows(results, dta_hits)
  }

  return(results)
}

```

### List files with filename prefix

```{r}
volume_path <- "/Volumes/dhs_indicator_dev/variables/chap07_fp"
# volume_path <- "/Volumes/{volume_catalog}/{volume_schema}/{volume_name}"
filename_prefix = c("NGIR", "ETIR", "NGMR", "ETMR")
dta_files <- list_dta_files(con, volume_path, filename_prefix)

print(dta_files)

```

### Get Recoded IR data from table using SQL interface

```{r}

query = "SELECT * from dhs_recode.et.etir71fl limit 10"
result <- DBI::dbGetQuery(con, query)
```

### Connect to your own cluster using Spark
Use Reticulate and databricks-connect package to connect 

```{r}
#| output: false
#| warning: false

library(here)
# inctall client that match your cluster
cluster_id = Sys.getenv("DATABRICKS_CLUSTER_ID")
py_path <- reticulate::install_python(version = "3.11")
venv_path <- here(".venv")
reticulate::virtualenv_create(envname = venv_path, python = py_path)
reticulate::use_virtualenv(venv_path)
Sys.setenv(UV_PYTHON_EXECUTABLE = file.path(venv_path, "Scripts", "python.exe"))
reticulate::py_install("databricks-connect==16.4")
# pysparklyr::install_databricks(cluster_id = cluster_id) 
reticulate::py_module_available("databricks.connect") 
```

### Connect to databricks cluster
```{r}
#| output: false
#| warning: false

library(sparklyr)
library(dplyr)
library(dbplyr)

spark_disconnect_all()
sc <- sparklyr::spark_connect(
  master     = paste0("https://",Sys.getenv("DATABRICKS_HOST")),
  cluster_id = Sys.getenv("DATABRICKS_CLUSTER_ID"),
  token      = Sys.getenv("DATABRICKS_TOKEN"),
  method     = "databricks_connect",
)

```

### Use Spark to Query / analyze DHS data if size is too large
```{r}
library(sparklyr)
library(dplyr)
library(DBI)

# Analysis of age_at_first_birth by region, education_level
df_ir <- sdf_sql(sc, sql("SELECT v511, v106, v024 FROM dhs_recode.ia.iair74fl"))
row_count <- df_ir %>% 
  summarise(count = n()) %>% 
  pull(count)
print(paste0("processing ", row_count, " rows"))
result <- df_ir %>%
  filter(!is.na(v511), !is.na(v106), !is.na(v024)) %>%
  group_by(region = v024, education = v106) %>%
  summarise(age_at_first_birth = mean(v511, na.rm = TRUE), .groups = "drop")

# Collect results into R
result_r <- collect(result)
print(result_r)

```
