---
title: "DHS data plot example"
date: "2024-11-11"
author: "YL"
output: html
execute:
  echo: true
---

# This is a simple DHS Analysis example


This is a simple example of how to create a bar plot from curated data showing patterns of family planning in R using ggplot2.

## Install packages
```{r results='hide', warning=FALSE, message=FALSE}
install.packages("dplyr", repos = "https://cloud.r-project.org/")
install.packages("readxl", repos = "https://cloud.r-project.org/")
install.packages("ggplot2", repos = "https://cloud.r-project.org/")
install.packages("ggpubr", repos = "https://cloud.r-project.org/")
library(dplyr)
library(readxl)
library(ggplot2)
library(ggpubr)
```
## Load curated data
```{r}
df <- read_excel("DHS_nonuse_other.xlsx")
head(df)
```
About the curated data:

The coefficients (Odds ratio, with confidence intervals) came from logistic regression model within each country. 
Each model controls for covariates including women's age, education, place of residence.

```{r, error=FALSE, warning=FALSE}
df$survey <-paste0(as.character(df$country)," ", as.character(df$year))
class(df$prevalence)
#create a binary indicating whether 
df$Significance <- ifelse(df$CI_low<1, "Not significant", "Significant")
table(df$Significance) ## 1/3 is significant 
prop.table(table(df$Significance))

### The prevalence of each survey, low -> high % 
p1 <- ggplot(df, aes(x=reorder(survey, -prevalence), y=prevalence)) +
  geom_bar(stat='identity') +
  coord_flip() +
  geom_hline(yintercept=5, color="red") +
  geom_vline(xintercept=43, color="red") 
```

display the plot 

```{r, fig.height=10, fig.width=8}
p1
```
