---
title: "Simple R Example with Map and Video"
date: "2024-09-23"
output: html
execute:
  echo: true
---

# This is a very basic quarto document for sharing your code


This is a simple example of how to create a bar plot in R using ggplot2.


```{r}
# Load necessary library
library(ggplot2)

# Create a simple dataset
data <- data.frame(
  category = c("A", "B", "C", "D"),
  values = c(3, 7, 9, 4)
)

# Create a bar plot
ggplot(data, aes(x = category, y = values)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  labs(title = "Simple Bar Plot", x = "Category", y = "Values")
```

::: {.callout-note}
You can create a callout note by using the `callout-note` class.
:::

Show some data your map

```{r}
# Load required packages
library(leaflet)
library(leaflet.extras) # Provides heatmap functionality

# Define the latitude and longitude for Seattle, WA
seattle_coords <- c(47.6062, -122.3321)

# Generate some sample data around the Seattle area
set.seed(123) # for reproducibility
lat <- rnorm(100, mean = 47.6062, sd = 0.01)  # latitude around Seattle
lng <- rnorm(100, mean = -122.3321, sd = 0.01) # longitude around Seattle
intensity <- rnorm(100, mean = 1, sd = 0.5)    # sample intensity values

# Create a data frame for heatmap
heatmap_data <- data.frame(lat, lng, intensity)

# Create the Leaflet map
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap tiles
  setView(lng = seattle_coords[2], lat = seattle_coords[1], zoom = 12) %>%
  addHeatmap(
    lng = heatmap_data$lng,
    lat = heatmap_data$lat,
    intensity = heatmap_data$intensity,
    blur = 20, max = 1, radius = 15
  )
```


::: {.callout-tip}
You can create callout-tips! Just use the `callout-tip` class.
:::


Quarto supports LaTeX-style math for rendering mathematical equations in documents. This type of math can be used to display both inline math expressions and block (display) math expressions. For example, the equation $E=mc^2$ is an example of inline math, while the equation $$ A = \pi r^2 $$ is the example of block math.
 

You can use mermaid to create diagrams in your documents. For example, the following code creates a simple flowchart:
```{mermaid}
graph TD;
  A[Start] --> B[Step 1];
  B --> C[Step 2];
  C --> D[End];

```

If you want to learn more, check this video:
{{< video https://youtu.be/_f3latmOhew?si=HE6bpTObWfmnAZzM >}}
